Supplement to: Learning to Infer Graphics Programs from Hand-Drawn Images

### 2018-11-17 poster.tex
progress on poster

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{76in} % A0 width: 46.8in
\setlength{\paperheight}{41in} % NIPS16 : 36
\setlength{\sepwid}{0.024\paperwidth} % Separation width (white space) between columns
\setlength{\onecolwid}{0.22\paperwidth} % Width of one column
\setlength{\twocolwid}{0.464\paperwidth} % Width of two columns
\setlength{\threecolwid}{0.708\paperwidth} % Width of three columns
\setlength{\topmargin}{-0.5in} % Reduce the top margin size
%-----------------------------------------------------------

@@

\author{Kevin Ellis, Daniel Ritchie, Lucas Morales, Mathias Sabl\'e Meyer,  Joshua B. Tenenbaum, Armando Solar-Lezama} % Author(s)

@@

\begin{block}{Learning from very few examples}
\begin{figure}[h]\centering
  \begin{minipage}{0.45\textwidth}\centering
    \includegraphics[width=\textwidth]{smallFractionSolving.png}
  \caption{Results averaged across 19 problems. Solid: \theSystem{} . Dashed: enumerating 100 programs. Dotted: MDL.}\label{flashPerformance}        \end{minipage}%
\hspace{0.025\textwidth}\begin{minipage}{0.45\textwidth}\centering
    \includegraphics[width=0.7\textwidth]{small_mdl.png}
    \caption{MDL learner vs \theSystem{} on one-shot learning. Predictions marginalize out the program.}\label{mdl}
  \end{minipage}
\end{figure}

%\textbf{Multiple explanations aids one-shot learning.} After a single example, the shortest program is probably not correct -- but one of the top sampled \emph{behaviors} (marginalizing over the program) might be.

\textbf{Good performance needs tilt correction.} w/o low-tilt approximation, get no samples for any of these problems after an hour, vs $\approx 1$ minute w/ \theSystem

\end{block}

%% \begin{block}{The Cost of Sampling}
%%   Amortized cost of 100 samples from \theSystem :

%%   \vspace{1cm}

%%   \hspace{9cm}\vspace{1cm}    \begin{tabular}{lll}\toprule
%%      1 example\hspace{0.9cm}      &  2 examples\hspace{0.9cm}          &     3 examples\\\midrule
%% $84 \pm 3$ sec &$21 \pm 1$ sec & $49\pm 3$ sec
%% \end{tabular}

%% \textbf{Good performance needs tilt correction.} w/o low-tilt approximation, get no samples for any of these problems after an hour!

%% \textbf{Can closely approximate posterior with little overhead.}

%% %\vspace{-2.5cm}
%% \begin{tabular}{cl}
%%   \begin{tabular}{c}
%%     \includegraphics[width=25cm]{trim-trade-off.png}
%%     \end{tabular}&
%%   \begin{tabular}{l}
%%     \parbox{15cm}{Accuracy (colored contours) vs Performance (monochrome cells) trade-off; upper bounds plotted. Performance measured in expected solver invocations; accuracy measured in log KL divergence.}
%%     \end{tabular}
%% \end{tabular}
%% %\end{figure}

%%   \end{block}
%----------------------------------------------------------------------------------------

@@

  \begin{alertblock}{NEW: Wake-Sleep DSL learning with \theSystem}

    \begin{minipage}[c]{0.45\textwidth}
    \begin{itemize}
    \item \textbf{Wake:} Synthesize programs from DSL + neural net (`recognition model') to guide search
    \item \textbf{Sleep-G:} Grow DSL, using grammar induction algorithms
    \item \textbf{Sleep-R:} Train neural net on programs found during waking (`experience replay') and samples from DSL (`dreaming')
    \end{itemize}
    \end{minipage}%
    \begin{minipage}[c]{0.48\textwidth}      
    \begin{tikzpicture}[scale=3,line width=1.5mm]
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $Q$};

    \draw [->] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
    \draw [->] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
    \draw [->] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);

    \draw [->,dashed] (DSL.north) to[out = 90,in = 180] node[fill=dblue!10,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
    \draw [->,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=dblue!10,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
  \end{scope}
    \end{tikzpicture}
      \end{minipage}
  \end{alertblock}

\begin{block}{\theSystem~outputs for three different task domains}

\hspace{-2cm}\includegraphics[width = 1.15\textwidth]{figures/3DSL.png}
{\small Top: Tasks from three domains we apply our algorithm to, each followed by the programs \theSystem~ discovers for them. Bottom: Several examples from learned DSL. Notice that learned DSL primitives can call each other, and that \theSystem~rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}
\end{block}

%----------------------------------------------------------------------------------------
%	ADDITIONAL INFORMATION
%----------------------------------------------------------------------------------------

\begin{block}{Bayesian framing}

  \begin{minipage}[c]{0.45\textwidth}
      \begin{tikzpicture}[scale=4,line width=1.5mm]

  \node[latent,scale=3] at (3.5,3) (dx){$\mathcal{D}$};
  \node[latent,scale=3] at ([xshift=2cm]dx) (zp){$p_n$};
  \node[obs,scale=3] at ([xshift=2cm]zp) (xp) {$x_n$};
  \plate {}{(zp)(xp)}{};
%  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.east) {(b) \\recognition model};
  \draw [->,red] (xp.south) to[out = -90,in = -90] node(nn){} (zp.south);
%  \draw [->,red] (tx.west) -- (zp.east);
  \draw [->] (dx.east) -- (zp.west);
  \draw [->] (zp.east) -- (xp.west);

  \node at (nn) {
    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.5,shift={+(-1,2.5)}}]
      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
      \node[rectangle] at (1,1) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (I-\name) at (0,-\y) {};
      \foreach \name / \y in {1,...,3}
          \node[neuron] (H-\name) at (1,-\y-0.5) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (O-\name) at (2,-\y) {};
      \foreach \source in {1,...,4}
          \foreach \dest in {1,...,3}
              \draw [-latex] (I-\source) -- (H-\dest);
      \foreach \source in {1,...,3}
          \foreach \dest in {1,...,4}
              \draw [-latex] (H-\source) -- (O-\dest);
    \end{tikzpicture}
  };
  \node[shift={+(0,-1.7)}] at (nn) { $Q$  };

      \end{tikzpicture}
      $x_n$: program synthesis problem. $p_n$: (latent) program solving $x_n$. $Q(p|x)$: (learned) neural recognition model. $\mathcal{D}$: (latent) DSL.
  \end{minipage}\qquad%
  \begin{minipage}[c]{0.4\textwidth}

  $$
\underbrace{p_n^* =   \argmax_{p_n}\probability[x_n|p_n]\probability[p_n|\mathcal{D}^*]}_{\text{\textbf{Wake}}}
$$
\vspace{0.5cm}
$$
\underbrace{\mathcal{D}^* = \argmax_{\mathcal{D}}    \probability[\mathcal{D}]\prod_n\sum_{p_n}\probability[x_n|p_n]\probability[p_n|\mathcal{D}]}_{\textbf{\text{Sleep-G}}}
$$
\vspace{0.5cm}
$$\underbrace{Q^* = \argmin_Q \text{KL}\left(\probability[p|x,\mathcal{D}] || Q(p|x)\right) }_{\text{\textbf{Sleep-R}}}$$
    \end{minipage}

\end{block}

>>>>

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{96in} % A0 width: 46.8in
\setlength{\paperheight}{48in} % NIPS16 : 36
\setlength{\sepwid}{0.024\paperwidth} % Separation width (white space) between columns
\setlength{\onecolwid}{0.22\paperwidth} % Width of one column
\setlength{\twocolwid}{0.464\paperwidth} % Width of two columns
\setlength{\threecolwid}{0.708\paperwidth} % Width of three columns
\setlength{\topmargin}{-0.5in} % Reduce the top margin size
%-----------------------------------------------------------

@@

\author{Kevin Ellis, Daniel Ritchie, Armando Solar-Lezama,  Joshua B. Tenenbaum} % Author(s)

@@

  \begin{block}{Error correction}
    `Top down' influences upon perception: reasoning engine (program synthesizer) can influence
    agent's percept through higher-level considerations like symmetry and alignment.
    \begin{align*}
      \hat{S}(I) &= \argmax_{S\in \mathcal{F}(I)} L_{\text{learned}}(I | \text{render}(S))\times \probability_\theta[S|I] \times\probability_{\beta} [ \text{program}(S)] \\
        \beta^* &= \argmax_{\beta} \expect \left[ \log \frac{\probability_{\beta} [\text{program}(S)]\times L_{\text{learned}}(I|\text{render}(S))\times \probability_\theta[S|I]}{\sum_{S'\in \mathcal{F}(I)} \probability_{\beta} [\text{program}(S')]\times L_{\text{learned}}(I|\text{render}(S'))\times \probability_\theta[S'|I]} \right]
    \end{align*}
    where $\mathcal{F}(I)$ is set of parses output by neural net on image $I$;
    $\probability_\beta[\cdot ]$ is prior over programs parameterized by $\beta$;
    and expectation is taken over a corpus of
    program synthesis problems.

    \vspace{2cm}
    \begin{figure}
        \includegraphics[width = 0.6\textwidth]{figures/programSuccess7.png}\hfill
  \includegraphics[width = 0.6\textwidth]{figures/programSuccess16.png}
  \caption{Left: hand drawings. Center: interpretations favored by the deep network. Right: interpretations favored after learning a prior over programs. The prior favors  simpler programs, thus (top) continuing the pattern of not having an arrow is preferred, or (bottom) continuing the ``binary search tree'' is preferred.}
      \end{figure}
    \end{block}

@@

### 2018-11-17 supplement.tex
progress on poster

\author{
Kevin Ellis\\
  MIT\\
  \texttt{ellisk@mit.edu} \\
  \And
  Daniel Ritchie\\
Brown University\\
 \texttt{daniel\_ritchie@brown.edu} \\
 \And
 Armando Solar-Lezama\\
 MIT\\
\texttt{asolar@csail.mit.edu} \\
\And
Joshua B. Tenenbaum \\
MIT\\
\texttt{jbt@mit.edu}
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

@@

\section{Learning a search policy}
\subsection{Modeling}\label{searchPolicyModel}
Recall from the main paper that our goal is to estimate the policy minimizing the following loss:
\begin{align}
\textsc{Loss}(\theta ; \mathcal{D})& =  \expect_{S\sim\mathcal{D}}\left[ \min_{\sigma\in \text{\textsc{Best}}(S)}\frac{t(\sigma | S)}{\pi_\theta (\sigma | S)}\right] + \lambda \Vert\theta\Vert_2^2\label{policyLoss}\\
\text{where }  \sigma \in\text{\textsc{Best}}(S) &\text{ if  a minimum cost program for }S \text{ is in }\sigma .\nonumber %\text{cost}(\text{program}(p)) = \min_{p\in \sigma \text{, s.t. }p\text{ evaluates to }S} \text{cost}(p) \nonumber
\end{align}
We make this optimization problem tractable by annealing our loss function during gradient descent:
\begin{align}
&  \textsc{Loss}_\beta(\theta ; \mathcal{D}) =  \expect_{S\sim\mathcal{D}}\left[ \textsc{SoftMinimum}_\beta \left\{
    \frac{t(\sigma | S)}{\pi_\theta (\sigma | S)}    : \sigma\in \text{\textsc{Best}}(S)\right\}\right] + \lambda \Vert\theta\Vert_2^2\label{softenedObjective}\\
  \text{where }&\textsc{SoftMinimum}_\beta (x_1,x_2,x_3,\cdots ) = \sum_n x_n  \frac{e^{-\beta x_n}}{\sum_{n'}e^{-\beta x_{n'}}}
\end{align}
Notice that $\textsc{SoftMinimum}_{\beta = \infty}(\cdot )$ is just $\min(\cdot )$.
We set the regularization coefficient $\lambda = 0.1$ and minimize equation~\ref{softenedObjective}
using Adam for 2000 steps, linearly increasing $\beta$ from $1$ to $2$.

>>>>

\author{
Kevin Ellis\\
  MIT\\
  \texttt{ellisk@mit.edu} \\
  \And
  Daniel Ritchie\\
Brown University\\
 \texttt{daniel\_ritchie@brown.edu} \\
 \AND
 Armando Solar-Lezama\\
 MIT\\
\texttt{asolar@csail.mit.edu} \\
\And
Joshua B. Tenenbaum \\
MIT\\
\texttt{jbt@mit.edu}
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

@@

\section{Learning a search policy}

Figure~\ref{yellow} provides additional intuition for
how the policy biases and informs the program synthesizer:
the entire program search space is carved up into smaller subsets,
and we search within each of these subsets \emph{simultaneously and in parallel},
but where the fraction of
compute time allocated to each subset
is proportional to the weight assigned to it by the policy.

\begin{figure}\centering
  \begin{tikzpicture}
    \node[anchor = west] at (0,5.25) {Entire program search space};
    \draw[fill = yellow,fill opacity = 0.15,draw = black] (0,0) rectangle (5,5);
    \draw [fill = yellow, opacity = 0.2] (0,0)--(0,2)--(4,0)--(0,0);
    \draw [fill = yellow, opacity = 0.4] (0,5)--(0,1)--(3,5);
    \draw [fill = yellow, opacity = 0.6] (2.5,5)--(3.2,0)--(5,0)--(5,5);
    \draw (0,2) -- node[below,sloped]{short programs} (4,0);
    \draw (0,2) -- node[above,sloped]{long programs} (4,0);
    \draw (2.5,5) -- node[above,sloped]{programs w/ reflections} (3.2,0);
    \draw (0,1) -- node[above,sloped]{ programs w/ loops} (3,5);
    \node(p1)[anchor =west ] at (5.5,3.5) {$\pi_\theta(\text{short, no loop/reflect}|S) = $};
    \draw [fill = yellow, fill opacity = 0.2,draw = black]  ([xshift = 0cm,yshift = -0.2cm]p1.east) rectangle ([xshift = 0.4cm,yshift = 0.2cm]p1.east);
    \node(p2)[anchor =west ] at ([yshift = -0.5cm]p1.west) {$\pi_\theta(\text{long, loops}|S) = $};
    \draw [fill = yellow, fill opacity = 0.4,draw = black]  ([xshift = 0cm,yshift = -0.2cm]p2.east) rectangle ([xshift = 0.4cm,yshift = 0.2cm]p2.east);
    \node(p3)[anchor =west ] at ([yshift = -0.5cm]p2.west) {$\pi_\theta(\text{long, no loop/reflect}|S) = $};
    \draw [fill = yellow, fill opacity = 0.15,draw = black]  ([xshift = 0cm,yshift = -0.2cm]p3.east) rectangle ([xshift = 0.4cm,yshift = 0.2cm]p3.east);
    \node(p4)[anchor =west ] at ([yshift = -0.5cm]p3.west) {$\pi_\theta(\text{long, reflects}|S) = $};
    \draw [fill = yellow, fill opacity = 0.6,draw = black]  ([xshift = 0cm,yshift = -0.2cm]p4.east) rectangle ([xshift = 0.4cm,yshift = 0.2cm]p4.east);
    \node(p5)[anchor =west ] at ([yshift = -0.5cm]p4.west) {\emph{etc.}};
  \end{tikzpicture}
  \caption{The bias-optimal search algorithm divides the entire (intractable) program search space in to (tractable) program subspaces (written $\sigma $), each of which contains a restricted set of programs. For example, one subspace might be short programs which don't loop. The policy $\pi$ predicts a distribution over program subspaces. The weight that $\pi$ assigns to a subspace is indicated by its yellow shading in the above figure, and is conditioned on the spec $S$.}\label{yellow}
  \end{figure}

\subsection{Modeling}\label{searchPolicyModel}
Recall from the main paper that our goal is to estimate the policy minimizing the following loss:
\begin{align}
\textsc{Loss}(\theta ; \mathcal{D})& =  \expect_{S\sim\mathcal{D}}\left[ \min_{\sigma\in \text{\textsc{Best}}(S)}\frac{t(\sigma | S)}{\pi_\theta (\sigma | S)}\right] + \lambda \Vert\theta\Vert_2^2\label{policyLoss}\\
\text{where }  \sigma \in\text{\textsc{Best}}(S) &\text{ if  a minimum cost program for }S \text{ is in }\sigma .\nonumber %\text{cost}(\text{program}(p)) = \min_{p\in \sigma \text{, s.t. }p\text{ evaluates to }S} \text{cost}(p) \nonumber
\end{align}
We make this optimization problem tractable by annealing our loss function during gradient descent:
\begin{align}
&  \textsc{Loss}_\beta(\theta ; \mathcal{D}) =  \expect_{S\sim\mathcal{D}}\left[ \textsc{SoftMinimum}_\beta \left\{
    \frac{t(\sigma | S)}{\pi_\theta (\sigma | S)}    : \sigma\in \text{\textsc{Best}}(S)\right\}\right] + \lambda \Vert\theta\Vert_2^2\label{softenedObjective}\\
  \text{where }&\textsc{SoftMinimum}_\beta (x_1,x_2,x_3,\cdots ) = \sum_n x_n  \frac{e^{-\beta x_n}}{\sum_{n'}e^{-\beta x_{n'}}}
\end{align}
Notice that $\textsc{SoftMinimum}_{\beta = \infty}(\cdot )$ is just $\min(\cdot )$.
We set the regularization coefficient $\lambda = 0.1$ and minimize equation~\ref{softenedObjective}
using Adam for 2000 steps, linearly increasing $\beta$ from $1$ to $2$.
